---
title: "DACSS 604 Final Project"
subtitle: "In Patterns to Problems lies hope for Solutions" 
author: "Theresa Szczepanski"
format: 
  revealjs:
    incremental: true
    #theme: sky
    logo: heron.jpg
    institute: "Rising Tide Charter Public School"
    footer: "DACSS 604 Spring 2025"
---

## My Audience

- ***Who***: Stakeholders at Rising Tide Charter Public School
    - Head of School, Principal, Assistant Principals, Department Chairs, and Board of Trustees
    - Make decisions on allocation of resources and evaluation of teaching staff
    - Write schoolwide and department wide goals for each school year
    - ***Very Limited Time***: School leaders are notoriously stretched thin with time
    
## My Audience (cont.)

- ***Technical Skill***
  - Some may have some statistical training but most do not
  - Assume general, college-educated audience level of understanding
  - Can assume familiarity with the concept of assessment and accountability testing in the K-12 public school sector.



## What are the [Star Screening Assessments](https://renaissance.widen.net/view/pdf/y7ouwvffec/R53639.pdf?t.download=true&u=zceria)?

- Computerized adaptive tests
- tested on national samples with cross-validation
- found to be ***reliable*** and ***valid predictors*** of students’ academic ***growth*** in ***literacy*** and ***mathematics*** with “convincing evidence”^[Source: [National Center for Intensive Interventions](https://charts.intensiveintervention.org/ascreening)]  




## How were they administered?

- ***Who***: Students in Grades 5-10

- ***What***: Star Literacy and Star Math

- ***When***: Three screenings (Fall, Winter, Spring) beginning in SY2023-2024


## What Metrics Do they Provide?

::: {.panel-tabset}

### Achievement

- Achievement Percentile (***PR***) 
  
    - Continuous, with minimum value of 1 and maximum value of 99
    - A student with a ***PR*** of 54 scored ***better*** than ***54%*** of students in their grade level.

    - A student is classified as ***At/Above Benchmark*** if their achievement percentile is ***at least 40***.
        
### Growth

- Student Growth Percentile (***SGP***) 
    - Measure of a student's improvement given their ***achievement*** in their first test.
    - Continuous with minimum value of 1 and maximum value of 99
    - A student with an ***SGP*** of 54 ***improved more*** than ***54%*** of students in their grade level who ***had similar achievement*** on their first screening test.

:::


   
    


- Growth








```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| label: setup
#| warning: false
#| message: false

# load the needed libraries

library(tidyverse)
library(ggplot2)
library(lubridate)
library(readxl)
library(hrbrthemes)
library(viridis)
library(ggpubr)
library(purrr)
library(plotly)
library(dplyr)
library(RColorBrewer)
library(stargazer)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
#source('readin_lib.R')
#source('student_perf_lib.R')

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read in Data Sets

star_dem_all_private<-read_csv("data_604/student_star_dem_all_private.csv")%>%
  mutate(`Grade` = as.factor(`Grade`))%>%
  mutate(`Grade` = recode_factor(`Grade`,
                                              "5" = "5",
                                              "6" = "6",
                                              "7" = "7",
                                              "8"= "8",
                                              "9"= "9",
                                              "10"= "10",
                                              .ordered = TRUE))%>%
  mutate(School_Year = as.factor(School_Year))%>%
   mutate(School_Year = recode_factor(School_Year,
                              "2023-2024" = "2023-2024",
                              "2024-2025" = "2024-2025",
                              .ordered = TRUE))%>%
  mutate(`Assignment Type` = as.factor(`Assignment Type`))%>%
  mutate(`Test 1 Benchmark Category` = recode_factor(`Test 1 Benchmark Category`,
                                              "At/Above Benchmark" = "At/Above Benchmark",
                                              "On Watch" = "On Watch",
                                              "Intervention" = "Intervention",
                                              "Urgent Intervention"= "Urgent Intervention",
                                              .ordered = TRUE))%>%
  mutate(IEP_Status = as.factor(IEP_Status))%>%
  mutate(`504_Status` = as.factor(`504_Status`))%>%
  mutate(`Low_Income_Status` = as.factor(Low_Income_Status))%>%
  mutate(`Gender` = as.factor(Gender) )%>%
  mutate(`Gender_Status` = case_when(
    Gender == "Male" ~ 0,
    Gender == "Female" ~ 1
  ))

#star_dem_all_private

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Historic Achievement Growth Summary
HistSchoolBenchSum<-read_csv("data_604/StarBenchmarkSummary_School.csv")%>% 
  mutate(Year = as.factor(Year))%>%
  mutate(`Screening Window` = as.factor(`Screening Window`))%>%
  mutate(`Benchmark_Status` = recode_factor(`Benchmark_Status`,
                                              "At/Above Benchmark" = "At/Above Benchmark",
                                              "Below Benchmark" = "Below Benchmark",
                                              .ordered = TRUE))%>%
  select(Year, `Screening Window`, Subject, `Benchmark_Status`, `% Students`)

#HistSchoolBenchSum
  

HistGradeGrowthSum<-read_csv("data_604/StarBenchmarkSummary_SGP_Sum.csv")%>%
  mutate(Year = as.factor(Year))%>%
  mutate(`Screening Window` = as.factor(`Screening Window`))%>%
  mutate(`Subject` = as.factor(`Subject`))%>%
  mutate(`Grade` = as.factor(Grade))%>%
  mutate(`Grade` = recode_factor(`Grade`,
                                              "Grade 5" = "Grade 5",
                                              "Grade 6" = "Grade 6",
                                              "Grade 7" = "Grade 7",
                                              "Grade 8"= "Grade 8",
                                              "Grade 9"= "Grade 9",
                                              "Grade 10"= "Grade 10",
                                              .ordered = TRUE))%>%
  select(Year, `Screening Window`, `Grade`, Subject, `Median SGP`)


#HistGradeGrowthSum

star_dem_all_private<-star_dem_all_private%>%
  mutate(Benchmark_Status = 
           case_when(`Test 1 PR` >= 40 ~ "At/Above Benchmark",
                     `Test 1 PR` < 40 ~ "Below Benchmark"))%>%
  mutate(Benchmark_Status = as.factor(Benchmark_Status))

```

## Research Question 1

-   Get in bed
-   Count sheep

## Are we meeting our Accountability Goal?


```{r, echo=FALSE, warning=FALSE, message=FALSE}
Bench_Sum_Plot<-HistSchoolBenchSum%>%
  filter(`Year` == "2024-2025" & `Screening Window` == "Winter" | `Year` == "2023-2024" & `Screening Window` == "Spring")%>%
  filter(`Benchmark_Status` == "At/Above Benchmark")%>%
  ggplot(aes(x=`Benchmark_Status`,fill = `Year`, y = `% Students`)) +
  geom_bar(position="dodge", stat = "identity") +
    geom_text(aes( y = `% Students`, label = `% Students`,
                     vjust = -.25), position = position_dodge(width = 1)) +

          #theme_classic() +
  theme(axis.title.x=element_blank(),
       #  axis.text.x = element_text(angle = 45, , hjust = 1),
          axis.ticks.x=element_blank())+
    scale_fill_brewer(palette = "Paired")+
    labs(
      y = "% Students",
       x= "Benchmark Category",
      title = "Rising Tide Benchmark Achievement by School Year",
      caption = "Source: Renaissance Star Literacy and Math Screening")+
  geom_hline(yintercept = 80, color = "grey")+
  annotate("text", x = .75, y = 85, label = "80% Goal") +
  facet_wrap(~Subject)
```

```{r, echo = FALSE}

#| eval: false
#| fig-cap: "Rising Tide Benchmark Achievement by School Year, Renaissance Star Literacy and Math"
#| fig-alt: >
#|    A bar blot with school year (ordered, factor) on the x-axis and the percentage of students
#|    classified as At/Above benchmark on the y-axis, faceted by subject (English Language Arts or Math)
#|    Plot illustrates the relationship between the three variables.

Bench_Sum_Plot

```


## Slide 4

-   Get in bed
-   Count sheep

## Slide 5

-   Get in bed
-   Count sheep

## Slide 6

-   Get in bed
-   Count sheep

## Slide 7

-   Get in bed
-   Count sheep

## Slide 8

-   Get in bed
-   Count sheep

## Slide 9

-   Get in bed
-   Count sheep

## Slide 10

-   Get in bed
-   Count sheep
