---
title: "DACSS 604 Final Project"
author: "Theresa Szczepanski"
format: 
 html:
    df-print: paged
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
    embed-resources: true
    self-contained-math: true
editor: 
  markdown: 
    wrap: 72
---

# Load Libraries

```{r}
#| label: setup
#| warning: false
#| message: false

# load the needed libraries

library(tidyverse)
library(ggplot2)
library(lubridate)
library(readxl)
library(hrbrthemes)
library(viridis)
library(ggpubr)
library(purrr)
library(plotly)
library(dplyr)
library(RColorBrewer)
library(stargazer)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
#source('readin_lib.R')
#source('student_perf_lib.R')

```

# Load the Data

## State Standards Mastery Summary

```{r}
ELA_Standards<-read_csv("data_604/State_Standards-_Mastery_Star_2425_ELA.csv")%>%
  mutate(Division = case_when(
    Grade == 9 | Grade == 10 ~ "HS",
    Grade == 5 | Grade == 6 | Grade == 7 | Grade == 8 ~ "MS"  
  ))%>%
  mutate(Grade = as.factor(Grade))%>%
  
  mutate(`Reporting Category` = as.factor(`Reporting Category`))%>%
  mutate(`Domain` = as.factor(`Domain`))%>%
  mutate(`Division` = as.factor(`Division`))%>%
  mutate(`Screening Window` = as.factor(`Screening Window`))%>%
mutate(`Text Type` = as.factor(`Text Type`))
  

ELA_Standards


```

## Growth/Demographic Summary

```{r}
star_dem_all_private<-read_csv("data_604/student_star_dem_all_private.csv")%>%
  mutate(`Grade` = as.factor(`Grade`))%>%
  mutate(`Grade` = recode_factor(`Grade`,
                                              "5" = "5",
                                              "6" = "6",
                                              "7" = "7",
                                              "8"= "8",
                                              "9"= "9",
                                              "10"= "10",
                                              .ordered = TRUE))%>%
  mutate(School_Year = as.factor(School_Year))%>%
    mutate(School_Year_Fac = School_Year)%>%
   mutate(School_Year = recode_factor(School_Year,
                              "2023-2024" = "2023-2024",
                              "2024-2025" = "2024-2025",
                              .ordered = TRUE))%>%
  mutate(`Assignment Type` = as.factor(`Assignment Type`))%>%
  mutate(`Test 1 Benchmark Category` = recode_factor(`Test 1 Benchmark Category`,
                                              "At/Above Benchmark" = "At/Above Benchmark",
                                              "On Watch" = "On Watch",
                                              "Intervention" = "Intervention",
                                              "Urgent Intervention"= "Urgent Intervention",
                                              .ordered = TRUE))%>%
  mutate(IEP_Status = as.factor(IEP_Status))%>%
  mutate(`504_Status` = as.factor(`504_Status`))%>%
  mutate(`Low_Income_Status` = as.factor(Low_Income_Status))%>%
  mutate(`Gender` = as.factor(Gender) )%>%
  mutate(`Gender_Status` = case_when(
    Gender == "Male" ~ 0,
    Gender == "Female" ~ 1
  ))

star_dem_all_private

```

## Historical Summary Data Frames

```{r}
HistSchoolBenchSum<-read_csv("data_604/StarBenchmarkSummary_School.csv")%>% 
  mutate(Year = as.factor(Year))%>%
  mutate(`Screening Window` = as.factor(`Screening Window`))%>%
  mutate(`Benchmark_Status` = recode_factor(`Benchmark_Status`,
                                              "At/Above Benchmark" = "At/Above Benchmark",
                                              "Below Benchmark" = "Below Benchmark",
                                              .ordered = TRUE))%>%
  select(Year, `Screening Window`, Subject, `Benchmark_Status`, `% Students`)

HistSchoolBenchSum
  

HistGradeGrowthSum<-read_csv("data_604/StarBenchmarkSummary_SGP_Sum.csv")%>%
  mutate(Year = as.factor(Year))%>%
  mutate(`Screening Window` = as.factor(`Screening Window`))%>%
  mutate(`Subject` = as.factor(`Subject`))%>%
  mutate(`Grade_Fac` = as.factor(Grade))%>%
  mutate(`Grade` = as.factor(Grade))%>%
  mutate(`Grade` = recode_factor(`Grade`,
                                              "Grade 5" = "Grade 5",
                                              "Grade 6" = "Grade 6",
                                              "Grade 7" = "Grade 7",
                                              "Grade 8"= "Grade 8",
                                              "Grade 9"= "Grade 9",
                                              "Grade 10"= "Grade 10",
                                              .ordered = TRUE))%>%
  select(Year, `Screening Window`, `Grade`, Subject, `Median SGP`)


HistGradeGrowthSum

```





# RQ1:

Are Rising Tide students making progress toward our achievement
accountability goal: “80% of Rising Students will test as at or above
Benchmark in Reading on the Renaissance Star Reading and Star
Mathematics screening assessments.?

## Historical Benchmark_Status

```{r}
star_dem_all_private<-star_dem_all_private%>%
  mutate(Benchmark_Status = 
           case_when(`Test 1 PR` >= 40 ~ "At/Above Benchmark",
                     `Test 1 PR` < 40 ~ "Below Benchmark"))%>%
  mutate(Benchmark_Status = as.factor(Benchmark_Status))

  



```

### Statistical Test

#### Acheivement: Math

```{r}
star_dem_math<- star_dem_all_private%>%
  filter(`Assignment Type` == "Star Math")

star_dem_math
fit_Math_achieve = lm(`Test 1 PR` ~ School_Year_Fac + `IEP_Status` + `504_Status` + `Low_Income_Status`, data = star_dem_math)

summary(fit_Math_achieve)
AIC(fit_Math_achieve)
BIC(fit_Math_achieve)
#star_dem_math
```



#### Acheivement: ELA

```{r}
star_dem_ELA<- star_dem_all_private%>%
  filter(`Assignment Type` == "Star Reading")
fit_ELA_achieve = lm(`Test 1 PR` ~ School_Year_Fac + `IEP_Status`, data = star_dem_ELA)

summary(fit_ELA_achieve)
AIC(fit_ELA_achieve)
BIC(fit_ELA_achieve)
```
### Visual: RQ1

```{r}
Bench_Sum_Plot<-HistSchoolBenchSum%>%
  filter(`Year` == "2024-2025" & `Screening Window` == "Winter" | `Year` == "2023-2024" & `Screening Window` == "Spring")%>%
  filter(`Benchmark_Status` == "At/Above Benchmark")%>%
  ggplot(aes(x=`Benchmark_Status`,fill = `Year`, y = `% Students`)) +
  geom_bar(position="dodge", stat = "identity") +
    geom_text(aes( y = `% Students`, label = `% Students`,
                     vjust = -.25), position = position_dodge(width = 1)) +

          #theme_classic() +
  theme(axis.title.x=element_blank(),
       #  axis.text.x = element_text(angle = 45, , hjust = 1),
          axis.ticks.x=element_blank())+
    scale_fill_brewer(palette = "Paired")+
    labs(
      y = "% Students",
       x= "Benchmark Category",
      title = "Rising Tide Benchmark Achievement by School Year",
      caption = "Source: Renaissance Star Literacy and Math Screening")+
  geom_hline(yintercept = 80, color = "grey")+
  #annotate("text", x = .75, y = 85, label = "80% Goal")# +
  facet_wrap(~Subject)
```

```{r}

#| eval: false
#| fig-cap: "Rising Tide Benchmark Achievement by School Year, Renaissance Star Literacy and Math"
#| fig-alt: >
#|    A bar blot with school year (ordered, factor) on the x-axis and the percentage of students
#|    classified as At/Above benchmark on the y-axis, faceted by subject (English Language Arts or Math)
#|    Plot illustrates the relationship between the three variables.

Bench_Sum_Plot

```



# Subject Level Fall Winter Growth Comparison

## ELA Growth

```{r}

Growth_Sum_ELA<-star_dem_all_private%>%
  filter(`School_Year` == "2024-2025")%>%
  filter(`Assignment Type` == "Star Reading")%>%
  group_by(Grade)%>%
  summarize(median_SGP= median(`SGP (Expectation=50)`))

Growth_Sum_ELA_Plot<-ggplot(Growth_Sum_ELA, aes(x=`Grade`, y = `median_SGP`,fill="skyblue")) +
    geom_bar(position="dodge", stat = "identity", fill="skyblue") +
    #facet_wrap(~Grade)+
    geom_text(aes( y = `median_SGP`, label = `median_SGP`,
                   vjust = -.25))+
  theme_classic()+
    #scale_color_brewer(palette = "Blues")+
    theme(axis.title.x=element_blank(),
          #axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    labs(
      y = "Median SGP",
       x= "Grade Level",
      title = "Literacy Student Growth Percentile by Grade Level",
      caption = "Source: 2024-2025 Renaissance Star Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x = 1., y = 52, label = "50th Percentile")
```

```{r}
#| eval: false
#| fig-cap: "Rising Tide Literacy Growth by Grade Level, 2024-2025 Renaissance Star Literacy Screening"
#| fig-alt: >
#|    A bar blot with Grade level (ordered, factor) on the x-axis and the median student growth percentile 
#|    (SGP) in literacy on the y-axis
#|    Plot illustrates the relationship between the two variables.

Growth_Sum_ELA_Plot

```
### ELA Historical Growth

```{r}

Hist_Growth_Sum_ELA_Plot<-HistGradeGrowthSum%>%
  filter(Subject == "ELA")%>%
  filter(`Screening Window` == "Fall-Winter")%>%
  ggplot(aes(x=`Grade`,fill = "skyblue", y = `Median SGP`)) +
    geom_bar(position="dodge", stat = "identity", fill = "skyblue") +
    #facet_wrap(~Grade)+
    geom_text(aes( y = `Median SGP`, label = `Median SGP`,
                   vjust = -.25), position = position_dodge(width = 1))+
  #theme_classic()+
    #scale_fill_brewer(palette = "Blues")+
    theme(axis.title.x=element_blank(),
          #axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    labs(
      y = "Median SGP",
       x= "Grade Level",
      title = "Literacy Student Growth Percentile by Grade Level",
      caption = "Source: Renaissance Star Literacy Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x = 1.55, y = 52, label = "50th Percentile")+
  facet_wrap(~Year)

```

```{r}
#| eval: true
#| fig-cap: "Rising Tide Literacy Growth by Grade Level, school year 2024-2025 and 2023-2024, Renaissance    Star Literacy Screening"
#| 
#| fig-alt: >
#|    A bar blot with Grade level (ordered, factor) on the x-axis and the median student growth percentile 
#|    (SGP) in literacy on the y-axis, faceted by school year (ordered, factor).
#|    Plot illustrates the relationship between the three variables.
#|    
Hist_Growth_Sum_ELA_Plot
```

## Math Growth

```{r}

Growth_Sum_Math<-star_dem_all_private%>%
  filter(`Assignment Type` == "Star Math")%>%
  filter(`School_Year` == "2024-2025")%>%
  group_by(Grade)%>%
  summarize(median_SGP= median(`SGP (Expectation=50)`))

Growth_Sum_Math_Plot<-ggplot(Growth_Sum_Math, aes(x=`Grade`,fill = `Grade`, y = `median_SGP`, fill="skyblue")) +
    geom_bar(position="dodge", stat = "identity", fill="skyblue") +
    #facet_wrap(~Grade)+
    geom_text(aes( y = `median_SGP`, label = `median_SGP`,
                   vjust = -.25))+
  theme_classic()+
    scale_fill_brewer(palette = "Blues")+
    theme(axis.title.x=element_blank(),
           axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    labs(
      y = "Median SGP",
       x= "Grade Level",
      title = "Math Student Growth Percentile by Grade Level",
      caption = "Source: 2024-2025 Renaissance Star Math Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x = 1.05, y = 53, label = "50th Percentile")
```

```{r}
#| eval: false
#| fig-cap: "Rising Tide Math Growth by Grade Level, 2024-2025 Renaissance Star Literacy Screening"
#| fig-alt: >
#|    A bar blot with Grade level (ordered, factor) on the x-axis and the median student growth percentile 
#|    (SGP) in math on the y-axis
#|    Plot illustrates the relationship between the two variables.

Growth_Sum_Math_Plot


```

```{r}

Hist_Growth_Sum_Math_Plot<-HistGradeGrowthSum%>%
  filter(Subject == "Math")%>%
  filter(`Screening Window` == "Fall-Winter")%>%
  ggplot(aes(x=`Grade`,fill = "skyblue", y = `Median SGP`)) +
    geom_bar(position="dodge", stat = "identity", fill = "skyblue") +
    #facet_wrap(~Grade)+
    geom_text(aes( y = `Median SGP`, label = `Median SGP`,
                   vjust = -.25), position = position_dodge(width = 1))+
  #theme_classic()+
    #scale_fill_brewer(palette = "Blues")+
    theme(axis.title.x=element_blank(),
          #axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    labs(
      y = "Median SGP",
       x= "Grade Level",
      title = "Math Student Growth Percentile by Grade Level",
      caption = "Source: Renaissance Star Math Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x = 5.5, y = 48, label = "50th Percentile")+
  facet_wrap(~Year)


```

```{r}
#| eval: false
#| fig-cap: "Rising Tide Math Growth by Grade Level, school year 2024-2025 and 2023-2024, Renaissance    Star Math Screening"
#| 
#| fig-alt: >
#|    A bar blot with Grade level (ordered, factor) on the x-axis and the median student growth percentile 
#|    (SGP) in math on the y-axis, faceted by school year (ordered, factor).
#|    Plot illustrates the relationship between the three variables.
#|    
Hist_Growth_Sum_Math_Plot

```


# RQ2 Grade 5 Math

Is there a relationship between Rising Tide’s Grade 5 Mathematics
curriculum initiatives and Grade 5 student growth in mathematics?

## Summary Data Frames

```{r}

Math_IEP_Sum<-star_dem_math%>%
  mutate(IEP = case_when(IEP_Status == 1 ~ "Yes",
                         IEP_Status == 0 ~ "No"))%>%
  filter(`Assignment Type` == "Star Math")%>%
  group_by(Grade, `School_Year`, IEP)%>%
  summarize(`Median SGP` = median(`SGP (Expectation=50)`))

```
## Visual: Growth Distribution

```{r}
MathHistGrowth<- star_dem_math

G5MathHistGrowth<-star_dem_math%>%
  filter(Grade == "5")

G5_Hist_Year <- G5MathHistGrowth %>%
  ggplot( aes(x=`SGP (Expectation=50)`, fill=School_Year)) +
    geom_histogram( color="#e9ecef", alpha=0.6,binwidth=10, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    theme_classic() +
    labs(fill="")#+
 # facet_wrap(~Year)

G5_Hist_Year

G5_Hist_Year_Facet <- G5MathHistGrowth %>%
  ggplot( aes(x=`SGP (Expectation=50)`, fill=School_Year)) +
    geom_histogram( color="#e9ecef", alpha=0.6,binwidth=15, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    theme_classic() +
    labs(fill="")+
  labs( y = "",
            title = "Grade 5 Student Growth Distribution",
            x = "Student Growth Percentile (SGP)",
            caption = "Star Math Assessments")+
  facet_wrap(~School_Year)

G5_Hist_Year_Facet

Growth_Hist<-MathHistGrowth%>%
  ggplot( aes(x= `SGP (Expectation=50)`)) +
    #geom_histogram(alpha=0.6, binwidth = 15) +
  
  #ggplot(aes()+
           geom_histogram(alpha = .9, binwidth = 15)+
    #scale_fill_viridis(discrete=TRUE) +
    #scale_color_viridis(discrete=TRUE) +
  theme_classic()

Growth_Hist
           

```
## Statistical Tests

```{r}

#summary(MathHistGrowth)

G5MathHistGrowth<-G5MathHistGrowth%>%
  mutate(School_Year_Factor = case_when(
        School_Year == "2023-2024" ~ "2023-2024",
        School_Year == "2024-2025" ~ "2024-2025"))%>%
  mutate(School_Year_Factor = as.factor(School_Year_Factor))%>%
  mutate(School_Year_Bool = case_when(
    `School_Year` == "2023-2024" ~ 0,
    `School_Year` == "2024-2025" ~ 1
  ))



fit_G5Math_growth_dem = lm(`SGP (Expectation=50)` ~ School_Year_Factor + `IEP_Status` + `Low_Income_Status` + `504_Status` + `Test 1 PR`, data = G5MathHistGrowth)
summary(fit_G5Math_growth_dem)
AIC(fit_G5Math_growth_dem)
BIC(fit_G5Math_growth_dem)


```
```{r}
stargazer(fit_G5Math_growth_dem, type = "text") 


```

### Diagnosing Model Fit

```{r}
plot(fit_G5Math_growth_dem, which = 1:6)
cooks= 4/179
cooks #.0223

```

## Visual G5 Math Growth

```{r}

Hist_Growth_Sum_G5Math_Plot<-HistGradeGrowthSum%>%
  filter(Subject == "Math")%>%
  filter(`Screening Window` == "Fall-Winter")%>%
  ggplot(aes(x=`Grade`,fill = ifelse(Grade == "5", "A", "B"), y = `Median SGP`)) +
    geom_bar(position="dodge", stat = "identity", show.legend=FALSE) +
    scale_fill_manual(values = c(A= "skyblue", B = "grey"))+
    #facet_wrap(~Grade)+
    geom_text(aes( y = `Median SGP`, label = `Median SGP`,
                   vjust = -.25), position = position_dodge(width = 1))+
  #theme_classic()+
    #scale_fill_brewer(palette = "Blues")+
    theme(axis.title.x=element_blank(),
          #axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    labs(
      y = "Median SGP",
       x= "Grade Level",
      title = "Math Student Growth Percentile by Grade Level",
      caption = "Source: Renaissance Star Math Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  #annotate("text", x = 5.5, y = 48, label = "50th Percentile")+
  facet_wrap(~Year)
```

```{r}
#| eval: true
#| fig-cap: "Rising Tide Math Growth by Grade Level, school year 2024-2025 and 2023-2024, Renaissance    Star Math Screening"
#| 
#| fig-alt: >
#|    A bar blot with Grade level (ordered, factor) on the x-axis and the median student growth percentile 
#|    (SGP) in math on the y-axis, faceted by school year (ordered, factor).
#|    Plot illustrates the relationship between the three variables.
#|    
Hist_Growth_Sum_G5Math_Plot

```

```{r}
Hist_Growth_G5Math_Plot<-HistGradeGrowthSum%>%
  filter(Subject == "Math")%>%
  filter(Grade == "5")%>%
  filter(`Screening Window` == "Fall-Winter")%>%
  ggplot(aes(x=`Year`,fill = "skyblue", y = `Median SGP`)) +
    geom_bar(position="dodge", stat = "identity", fill = "skyblue") +
    #facet_wrap(~Grade)+
    geom_text(aes( y = `Median SGP`, label = `Median SGP`, 
                   vjust = -.25), position = position_dodge(width = 1))+
  theme_classic()+
    scale_fill_brewer(palette = "Paired")+
    theme(axis.title.x=element_blank(),
          #axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    labs(
      y = "Median SGP",
       x= "School Year",
      title = "Improvement in Grade 5 Student Growth Percentile",
      caption = "Source: Renaissance Star Math Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x = 1., y = 52, label = "Goal 50th Percentile")#+
  #theme_classic()
  #facet_wrap(~Year)


```

```{r}
#| eval: true
#| fig-cap: "Rising Tide Grade 5 Math Growth by school year 2024-2025 and 2023-2024, Renaissance Star Math Screening"
#| 
#| fig-alt: >
#|    A bar blot with school year (ordered, factor) on the x-axis and the median G5 student growth  
#|    percentile (SGP) in math on the y-axis.
#|    Plot illustrates the relationship between the two variables.
#|    
Hist_Growth_G5Math_Plot

```


## Visual G5 IEP Growth

```{r}
 IEP_Growth_G5Math_Plot<- Math_IEP_Sum%>%
  filter(Grade == "5")%>%
  ggplot(aes(x=`IEP`,fill = `School_Year`, y = `Median SGP`)) +
      geom_bar(position="dodge", stat = "identity") +
      geom_text(aes( y = `Median SGP`, x = `IEP`, label = `Median SGP`), position = position_dodge(width=.9),
                   vjust = 1)+
  theme_classic()+
      theme(#axis.title.x=element_blank(),
            #axis.text.x=element_blank(),
           # axis.text.x=element_text(angle =45, hjust = 1),
            axis.ticks.x=element_blank())+
      scale_fill_brewer(palette = "Paired")+
      
 # facet_wrap(~`Screening Window`)+
      labs(
        y = "Median SGP",
         x= "IEP Status",
        title = "Improved Growth For All Grade 5 Learners",
        caption = "Source: Renaissance Star Math Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x =.70, y = 52, label = "50th Percentile")#+
```

```{r}
#| eval: true
#| fig-cap: "Rising Tide Grade 5 Math Growth by disability status (IEP) and school year 2024-2025 and 2023-2024, Renaissance Star Math Screening"
#| 
#| fig-alt: >
#|    A clustered, bar blot with IEP status (ordered, factor) on the x-axis and the 
#|    median G5 student growthpercentile (SGP) in math on the y-axis.
#|    Bars are colored by the school year (ordered, factor)
#|    Plot illustrates the relationship between the three variables.
#| 

IEP_Growth_G5Math_Plot
```




# RQ 3: ELA State Standards Mastery

Is there a relationship between the text type of a literacy standard
(informational vs. literature) and Rising Tide High School students’
mastery of the standard?


##Visual: Distribution ELA Standards Mastery

```{r}

HS_ELA_Standards<-ELA_Standards%>%
  filter(Grade == 9|Grade == 10)%>%
  filter(`Text Type` == "Informational Text" | `Text Type` == "Literature")
HS_ELA_Standards



HS_ELA_Text_Facet <- HS_ELA_Standards %>%
  ggplot( aes(x=`% Secure`, fill=`Text Type`)) +
    geom_histogram( color="#e9ecef", alpha=0.6,binwidth=15, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    theme_classic() +
   
  labs( y = "",
            title = "High School State Standard Mastery Distribution",
            x = "% Students Secure",
            caption = "Star Reading: State Standards Mastery")+
  facet_wrap(~`Text Type`)

HS_ELA_Text_Facet



```

## Statistical Tests


```{r}
fit_ELA_text = lm(`% Secure` ~ ( `Text Type` + `Screening Window` + `Domain`), data = ELA_Standards)
summary(fit_ELA_text)

fit_HS_ELA_text_domain = lm(`% Secure` ~ ( `Text Type` + `Screening Window` + `Domain`), data = HS_ELA_Standards)
summary(fit_HS_ELA_text_domain)
AIC(fit_HS_ELA_text_domain)
BIC(fit_HS_ELA_text_domain)

```

```{r}

stargazer(fit_ELA_text, type = "text")
stargazer(fit_HS_ELA_text_domain, type = "text")

```


## Visual Mastery by Text Type

### SchoolWide

```{r}

#ELA_Standards

ELA_Text_Summary<-ELA_Standards%>%
  filter(`Reporting Category` == "Reading")%>%
   group_by(`Text Type`, `Screening Window`)%>%
  summarize(`Mean % Secure` = round(mean(`% Secure`)),
            `Median % Secure` = round(median(`% Secure`)))

#ELA_Text_Summary

ELA_Text_Summary%>%
  filter(`Screening Window` != "Spring")%>%
  ggplot(aes(x=`Text Type`,fill = `Text Type`, y = `Mean % Secure`)) +
      geom_bar(position="dodge", stat = "identity") +
      geom_text(aes( y = `Mean % Secure`, x = `Text Type`, label = `Mean % Secure`), position = position_dodge(width=.9),
                   vjust = 1)+
      theme(axis.title.x=element_blank(),
            #axis.text.x=element_blank(),
            axis.text.x=element_text(angle =45, hjust = 1),
            axis.ticks.x=element_blank())+
      scale_fill_brewer(palette = "Paired")+
      
  facet_wrap(~`Screening Window`)+
      labs(
        y = "Mean % Secure",
         x= "Domainl",
        title = "School State Standard Mastery by Text Type ELA",
        caption = "2024-2025 Star Reading Screening")


ELA_Text_Domain_Summary<-ELA_Standards%>%
  filter(`Reporting Category` == "Reading")%>%
   group_by(`Text Type`, `Domain`, `Screening Window`)%>%
  summarize(`Mean % Secure` = round(mean(`% Secure`)),
            `Median % Secure` = round(median(`% Secure`)))

#ELA_Text_Domain_Summary

ELA_Text_Div_Summary<-ELA_Standards%>%
  filter(`Reporting Category` == "Reading")%>%
   group_by(`Text Type`, `Division`, `Screening Window`)%>%
  summarize(`Mean % Secure` = round(mean(`% Secure`)),
            `Median % Secure` = round(median(`% Secure`)))

#ELA_Text_Div_Summary
```
### MS vs. HS
```{r}
ELA_Text_Div_Summary%>%
  filter(`Screening Window` != "Spring")%>%
  ggplot(aes(x=`Text Type`,fill = `Screening Window`, y = `Mean % Secure`)) +
      geom_bar(position="dodge", stat = "identity") +
      geom_text(aes( y = `Mean % Secure`, x = `Text Type`, label = `Mean % Secure`), position = position_dodge(width=.9),
                   vjust = 1)+
      theme(axis.title.x=element_blank(),
            #axis.text.x=element_blank(),
            axis.text.x=element_text(angle =45, hjust = 1),
            axis.ticks.x=element_blank())+
      scale_fill_brewer(palette = "Paired")+
      
  facet_wrap(~`Division`)+
      labs(
        y = "Mean % Secure",
         x= "Domainl",
        title = "School State Standard Mastery by Text Type ELA",
        caption = "2024-2025 Star Reading Screening")

```


### By Domain Cluster
```{r}

ELA_Text_Domain_Summary%>%
  filter(`Screening Window` != "Spring")%>%
  ggplot(aes(x=`Domain`,fill = `Text Type`, y = `Mean % Secure`)) +
      geom_bar(position="dodge", stat = "identity") +
      geom_text(aes( y = `Mean % Secure`, x = `Domain`, label = `Mean % Secure`), position = position_dodge(width=.9),
                   vjust = 1)+
      theme(axis.title.x=element_blank(),
            #axis.text.x=element_blank(),
            axis.text.x=element_text(angle =45, hjust = 1),
            axis.ticks.x=element_blank())+
      scale_fill_brewer(palette = "Paired")+
      
  facet_wrap(~`Screening Window`)+
      labs(
        y = "Mean % Secure",
         x= "Domainl",
        title = "School State Standard Mastery by Text Type ELA",
        caption = "2024-2025 Star Reading Screening")

```

## Visual: HS Informational Text vs. Literature Comparison

```{r}
ELA_Standards
ELA_TextType_Sum_HS<-ELA_Standards%>%
  filter(`Division` == "HS")%>%
  filter (`Screening Window` != "Spring")%>%
  filter(`Reporting Category` == "Reading")%>%
  group_by(`Screening Window`, Grade, `Text Type`)%>%
  summarize(`Median % Secure` = round(median(`% Secure`)))


 ELA_TextType_Sum_HS%>%
  ggplot(aes(x=`Text Type`,fill = `Screening Window`, y = `Median % Secure`)) +
  geom_bar(position="dodge", stat = "identity") +
  #scale_x_reverse()+
    geom_text(aes( y = `Median % Secure`, label = `Median % Secure`,
                     vjust = -.25), position = position_dodge(width = 1)) +

          #theme_classic() +
  theme(axis.title.x=element_blank(),
         # axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    scale_fill_brewer(palette = "Paired")+
    labs(
      y = "Median % Secure",
       x= "Text Type",
      title = "HS Reading Standards Mastery by Text Type",
      caption = paste("Fall-Winter Screening", "2024-2025"))+
   facet_wrap(~Grade)




```

## Visual: MS Informational Text vs. Literature Mastery: Fall-Winter

```{r}
ELA_TextType_Sum_MS<-ELA_Standards%>%
  filter(`Division` == "MS")%>%
  filter (`Screening Window` != "Spring")%>%
  filter(`Reporting Category` == "Reading")%>%
  group_by(`Screening Window`, Grade, `Text Type`)%>%
  summarize(`Median % Secure` = round(median(`% Secure`)))


 ELA_TextType_Sum_MS%>%
  ggplot(aes(x=`Text Type`,fill = `Screening Window`, y = `Median % Secure`)) +
  geom_bar(position="dodge", stat = "identity") +
  #scale_x_reverse()+
    geom_text(aes( y = `Median % Secure`, label = `Median % Secure`,
                     vjust = -.25), position = position_dodge(width = 1)) +

          #theme_classic() +
  theme(axis.title.x=element_blank(),
         # axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    scale_fill_brewer(palette = "Paired")+
    labs(
      y = "Median % Secure",
       x= "Text Type",
      title = "MS Reading Standards Mastery by Text Type",
      caption = paste("Fall-Winter Screening", "2024-2025"))+
   facet_wrap(~Grade)


```





















