---
title: "[A Rising Tide Lifts all Boats?]{style='color:navy'}"
subtitle: "Analysis of SY2024-2025 Grade 5 Math Initiative" 
author: "Theresa Szczepanski"
format: 
  revealjs:
    incremental: true
    #theme: sky
    logo: heron.jpg
    institute: "Rising Tide Charter Public School"
    footer: "DACSS 604 Spring 2025"
  
---

## [My Audience]{style="color:navy;"}


- ***Who***: Stakeholders at Rising Tide Charter Public School

    - Head of School, Principal, Assistant Principals, Department Chairs, and Board of Trustees
    
- ***What*** is the purpose?
    - Inform decisions on allocation of resources and evaluation of teaching staff and academic program
  

## [My Audience (cont.)]{style="color:navy;"}

- ***Technical Skill***
  - some may have some statistical training but most do not
  - assume general, college-educated audience level of understanding
  - can assume familiarity with the concept of assessment and accountability testing in the K-12 public school sector



## [What are the [Star Screening Tests](https://renaissance.widen.net/view/pdf/y7ouwvffec/R53639.pdf?t.download=true&u=zceria)?]{style="color:navy;"}

- computerized, adaptive tests
- tested on national samples with cross-validation
- found to be ***reliable*** and ***valid predictors*** of students’ academic ***growth*** in ***literacy*** and ***mathematics*** with “convincing evidence”^[Source: [National Center for Intensive Interventions](https://charts.intensiveintervention.org/ascreening)]  




## [How were they administered?]{style="color:navy;"}

- ***Who***: Students in Grades 5-10

- ***What***: Star Literacy and Star Math

- ***When***: Three screenings (Fall, Winter, Spring) beginning in SY2023-2024


    
## [What Metrics Do they Provide?]{style="color:navy;"}

- Test ***Achievement*** Percentile

- Student ***Growth*** Percentile

## [Test Achievement Percentile: ***PR***]{style="color:navy;"}

  
- Minimum value of 1 and maximum value of 99
- A student with a ***PR*** of 54 scored ***higher*** than ***54%*** of students in their grade level.

- A student is classified as ***At/Above Benchmark*** for their grade level if their achievement percentile is ***at least 40***.
        
## [Student Growth Percentile: ***SGP***]{style="color:navy;"}

- Measure of a student's ***improvement*** given their ***achievement*** on their first test.

- Minimum value of 1 and maximum value of 99

- A student with an ***SGP*** of 54 ***improved more*** than ***54%*** of students in their grade level who ***had similar achievement*** on their first screening test.



## [What question will we examine?]{style="color:navy;"}

- Is there a relationship between Rising Tide’s 2024-2025 Grade 5 Mathematics
programming initiatives and Grade 5 student growth in mathematics?


- ***Why?***
    - In the Spring of 2022, 2023, and 2024, Rising Tide Grade 5 math students' average growth percentiles on the MCAS were ***considerably*** below the state average of 50; ***75%*** of students were ***below*** the ***25th*** percentile for growth in 2023.^[[Source: Rising Tide 2023 Accountability Report](https://profiles.doe.mass.edu/mcas/achievement_level.aspx?linkid=32&orgcode=04830000&orgtypecode=5&&fycode=2022)]


## [What changed in SY2024-2025?]{style="color:navy;"}

- ***Schedule***: Increase instructional hours dedicated to math in Grade 5

- ***Instruction***: Train teachers on the use of the G5 Math curricular materials



## [Schedule]{style="color:navy;"}

<p style="font-size: 0.8em;">

  - Provide a year-long course calendar mapping of lessons to weekly school schedule
  
  - Move 4 units from the G5 Math course to the G5 STE[**M**]{style="color:blue;"}  course

</p>

## [Instruction]{style="color:navy;"}



  
  - Provide template for the use of time in a 55 minute block
  
      
  
  - Set clear expectations on frequency of assessment and use of team-written assessments
  
  
  - Provide bank of ***high quality, supplemental resources*** for numeracy
  
  - Provide district-level, standards-aligned assessments one month after each unit for analysis
  


  
    



```{r, echo=FALSE, warning=FALSE, message=FALSE}
#| label: setup
#| warning: false
#| message: false
# load the needed libraries
library(tidyverse)
library(ggplot2)
library(lubridate)
library(readxl)
library(hrbrthemes)
library(viridis)
library(ggpubr)
library(purrr)
library(plotly)
library(dplyr)
library(RColorBrewer)
library(stargazer)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
#source('readin_lib.R')
#source('student_perf_lib.R')

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read in Data Sets

star_dem_all_private<-read_csv("data_604/student_star_dem_all_private.csv")%>%
  mutate(`Grade` = as.factor(`Grade`))%>%
  mutate(`Grade` = recode_factor(`Grade`,
                                              "5" = "5",
                                              "6" = "6",
                                              "7" = "7",
                                              "8"= "8",
                                              "9"= "9",
                                              "10"= "10",
                                              .ordered = TRUE))%>%
  mutate(School_Year = as.factor(School_Year))%>%
   mutate(School_Year = recode_factor(School_Year,
                              "2023-2024" = "2023-2024",
                              "2024-2025" = "2024-2025",
                              .ordered = TRUE))%>%
  mutate(`Assignment Type` = as.factor(`Assignment Type`))%>%
  mutate(`Test 1 Benchmark Category` = recode_factor(`Test 1 Benchmark Category`,
                                              "At/Above Benchmark" = "At/Above Benchmark",
                                              "On Watch" = "On Watch",
                                              "Intervention" = "Intervention",
                                              "Urgent Intervention"= "Urgent Intervention",
                                              .ordered = TRUE))%>%
  mutate(IEP_Status = as.factor(IEP_Status))%>%
  mutate(`504_Status` = as.factor(`504_Status`))%>%
  mutate(`Low_Income_Status` = as.factor(Low_Income_Status))%>%
  mutate(`Gender` = as.factor(Gender) )%>%
  mutate(`Gender_Status` = case_when(
    Gender == "Male" ~ 0,
    Gender == "Female" ~ 1
  ))

#star_dem_all_private

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Historic Achievement Growth Summary
HistSchoolBenchSum<-read_csv("data_604/StarBenchmarkSummary_School.csv")%>% 
  mutate(Year = as.factor(Year))%>%
  mutate(`Screening Window` = as.factor(`Screening Window`))%>%
  mutate(`Benchmark_Status` = recode_factor(`Benchmark_Status`,
                                              "At/Above Benchmark" = "At/Above Benchmark",
                                              "Below Benchmark" = "Below Benchmark",
                                              .ordered = TRUE))%>%
  select(Year, `Screening Window`, Subject, `Benchmark_Status`, `% Students`)

#HistSchoolBenchSum
  

HistGradeGrowthSum<-read_csv("data_604/StarBenchmarkSummary_SGP_Sum.csv")%>%
  mutate(Year = as.factor(Year))%>%
  mutate(`Screening Window` = as.factor(`Screening Window`))%>%
  mutate(`Subject` = as.factor(`Subject`))%>%
  mutate(`Grade` = as.factor(Grade))%>%
  mutate(`Grade` = recode_factor(`Grade`,
                                              "Grade 5" = "Grade 5",
                                              "Grade 6" = "Grade 6",
                                              "Grade 7" = "Grade 7",
                                              "Grade 8"= "Grade 8",
                                              "Grade 9"= "Grade 9",
                                              "Grade 10"= "Grade 10",
                                              .ordered = TRUE))%>%
  select(Year, `Screening Window`, `Grade`, Subject, `Median SGP`)


#HistGradeGrowthSum

star_dem_all_private<-star_dem_all_private%>%
  mutate(Benchmark_Status = 
           case_when(`Test 1 PR` >= 40 ~ "At/Above Benchmark",
                     `Test 1 PR` < 40 ~ "Below Benchmark"))%>%
  mutate(Benchmark_Status = as.factor(Benchmark_Status))

```



## [What did we observe?]{style="color:navy;"}
::: aside
<p style="font-size: 0.7em;">
Grade 5 students with the same Test 1 achievement percentile, disability status, and income status are predicted to have an ***SGP 10 percentiles higher*** in SY2024-2025 than in SY2023-2024.
</p>
:::
```{r, echo=FALSE, warning=FALSE, message=FALSE}

star_dem_math<- star_dem_all_private%>%
  filter(`Assignment Type` == "Star Math")

Math_IEP_Sum<-star_dem_math%>%
  mutate(IEP = case_when(IEP_Status == 1 ~ "Yes",
                         IEP_Status == 0 ~ "No"))%>%
  filter(`Assignment Type` == "Star Math")%>%
  group_by(Grade, `School_Year`, IEP)%>%
  summarize(`Median SGP` = median(`SGP (Expectation=50)`))

#Math_IEP_Sum

Math_Gen_Sum<-star_dem_math%>%
  filter(`Assignment Type` == "Star Math")%>%
  group_by(Grade, `School_Year`, Gender)%>%
  summarize(`Median SGP` = median(`SGP (Expectation=50)`))

#Math_Gen_Sum

#star_dem_math

Math_Gen_IEP_Sum<-star_dem_math%>%
 filter(`Grade` == "5")%>%
  group_by(`School_Year`, Gender, IEP_Status)%>%
  summarize(`Median SGP` = median(`SGP (Expectation=50)`))

#Math_Gen_IEP_Sum

Math_Gen_IEP_Achieve_Sum<-star_dem_math%>%
 filter(`Grade` == "5")%>%
  group_by(`School_Year`, Gender, IEP_Status)%>%
  summarize(`Median Test 1 PR` = median(`Test 1 PR`))

#Math_Gen_IEP_Achieve_Sum

text_pos_data<-data.frame(
  x = c(3.85, 3.85),
  y = c(48, 48),
  cyl =c(4,6),
  label = c("50th Percentile", "")
)

```
```{r, echo=FALSE, warning=FALSE, message=FALSE}


Hist_Growth_Sum_G5Math_Plot<-HistGradeGrowthSum%>%
  filter(Subject == "Math")%>%
  filter(`Screening Window` == "Fall-Winter")%>%
  filter(`Grade` != "9")%>%
  filter(`Grade` != "10")%>%
  ggplot(aes(x=`Grade`,fill = ifelse(Grade == "5", "A", "B"), y = `Median SGP`)) +
    geom_bar(position="dodge", stat = "identity", show.legend=FALSE) +
    scale_fill_manual(values = c(A= "skyblue", B = "grey"))+
    #facet_wrap(~Grade)+
    geom_text(aes( y = `Median SGP`, label = `Median SGP`,
                   vjust = -.25), position = position_dodge(width = 1))+
  #theme_classic()+
    #scale_fill_brewer(palette = "Blues")+
    theme(axis.title.x=element_blank(),
          #axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    labs(
      y = "Median SGP",
       x= "Grade Level",
      title = "Math Student Growth Percentile by Grade Level",
      caption = "Source: Renaissance Star Math Screening")+
  geom_hline(yintercept = 50, color = "grey")+
  #geom_text(aes(x=x, y=y, label = label), data = text_pos_data, inherit.aes=FALSE)+
  #facet_wrap(~cyl)+
  #annotate("text", x = 3.85, y = 48, label = "50th Percentile")+
  facet_wrap(~Year)

Hist_Growth_Sum_G5Math_Plot2<-HistGradeGrowthSum%>%
  filter(Subject == "Math")%>%
  filter(`Screening Window` == "Fall-Winter")%>%
  filter(`Grade` == "5")%>%
  ggplot(aes(x=`Year`,fill = ifelse(Year == "2024-2025", "A", "B"), y = `Median SGP`)) +
    geom_bar(position="dodge", stat = "identity", show.legend=FALSE) +
    scale_fill_manual(values = c(A= "skyblue", B = "skyblue"))+
    #facet_wrap(~Grade)+
    geom_text(aes( y = `Median SGP`, label = `Median SGP`,
                   vjust = -.25), position = position_dodge(width = 1))+
  theme_classic()+
    #scale_fill_brewer(palette = "Blues")+
    theme(axis.title.x=element_blank(),
          #axis.text.x=element_blank(),
          axis.ticks.x=element_blank())+
    labs(
      y = "Median SGP",
       x= "School Year",
      title = "G5 Math Student Growth Percentile by School Year",
      caption = "Source: Renaissance Star Math Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x = 1., y = 48, label = "50th Percentile")
```

```{r, echo = FALSE}
#| eval: true
#| fig-alt: >
#|    A bar blot with Grade level (ordered, factor) on the x-axis and the median student growth percentile 
#|    (SGP) in math on the y-axis, faceted by school year (ordered, factor).
#|    Plot illustrates the relationship between the three variables.
#|    
Hist_Growth_Sum_G5Math_Plot

```
## [What did we observe (option 2)?]{style="color:navy;"}
::: aside
<p style="font-size: 0.7em;">
Grade 5 students with the same Test 1 achievement percentile, disability status, and income status are predicted to have an ***SGP 10 percentiles higher*** in SY2024-2025 than in SY2023-2024.
</p>
:::
```{r, echo = FALSE}
#| eval: true
#| 
#| fig-alt: >
#|    A bar blot with School year (ordered, factor) on the x-axis and the median 5th grade student growth percentile 
#|    (SGP) in math on the y-axis.
#|    Plot illustrates the relationship between the two variables.
#|    
Hist_Growth_Sum_G5Math_Plot2

```


## [What about our students on IEPs?]{style="color:navy;"}
::: aside
<p style="font-size: 0.7em;">
Grade 5 students on IEPs have a considerably lower median ***SGP*** than their non-disabled peers; however the difference between the median SGP of students on IEPs and their non-disabled peers was ***reduced*** by more than ***200%*** in SY2024-2025.
</p>
:::

```{r, echo=FALSE, warning=FALSE, message=FALSE}

 IEP_Growth_G5Math_Plot<- Math_IEP_Sum%>%
  filter(Grade == "5")%>%
  ggplot(aes(x=`School_Year`,fill = `IEP`, y = `Median SGP`)) +
      geom_bar(position="dodge", stat = "identity") +
      geom_text(aes( y = `Median SGP`, x = `School_Year`, label = `Median SGP`), position = position_dodge(width=.9),
                   vjust = 1)+
  theme_classic()+
      theme(#axis.title.x=element_blank(),
            #axis.text.x=element_blank(),
           # axis.text.x=element_text(angle =45, hjust = 1),
            axis.ticks.x=element_blank())+
      scale_fill_brewer(palette = "Paired")+
      
 # facet_wrap(~`Screening Window`)+
      labs(
        y = "Median SGP",
         x= "IEP Status",
        title = "G5 Student Growth Percentile by Year and IEP Status",
        caption = "Source: Renaissance Star Math Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x =.70, y = 52, label = "Goal 50th Percentile")#+
```


```{r, echo = FALSE}

#| eval: false
#| fig-cap: "Rising Tide Grade 5 Math Growth by disability status (IEP) and School Year, Renaissance Star MathScreening"
#| 
#| fig-alt: >
#|    A clustered, bar blot with School Year status (ordered, factor) on the x-axis and the 
#|    median G5 student growthpercentile (SGP) in math on the y-axis.
#|    Bars are colored by IEP_Status (ordered, factor)
#|    Plot illustrates the relationship between the three variables.
#|    
IEP_Growth_G5Math_Plot

```

## [Future Questions]{style="color:navy;"}

-   Examxine growth/acheivement as measure by MCAS Spring 2025 Preliminary results

```{r, echo=FALSE, warning=FALSE, message=FALSE}
Gen_Growth_G5Math_Plot<- Math_Gen_Sum%>%
  filter(Grade == "5")%>%
  ggplot(aes(x=`Gender`,fill = `School_Year`, y = `Median SGP`)) +
      geom_bar(position="dodge", stat = "identity") +
      geom_text(aes( y = `Median SGP`, x = `Gender`, label = `Median SGP`), position = position_dodge(width=.9),
                   vjust = 1)+
  theme_classic()+
      theme(axis.title.x=element_blank(),
            #axis.text.x=element_blank(),
           # axis.text.x=element_text(angle =45, hjust = 1),
            axis.ticks.x=element_blank())+
      scale_fill_brewer(palette = "Paired")+
      
 # facet_wrap(~`Screening Window`)+
      labs(
        y = "Median SGP",
         x= "Gender",
        title = "Improved Growth For All Grade 5 Learners",
        caption = "Source: Renaissance Star Screening")+
   geom_hline(yintercept = 50, color = "grey")+
  annotate("text", x =.70, y = 52, label = "Goal 50th Percentile")#+
  
  
```

```{r, echo = FALSE}

#| eval: false
#| fig-cap: "Rising Tide Grade 5 Math Growth by disability status (IEP) and school year 2024-2025 and 2023-2024, Renaissance Star Math Screening"
#| 
#| fig-alt: >
#|    A clustered, bar blot with Gender (ordered, factor) on the x-axis and the 
#|    median G5 student growthpercentile (SGP) in math on the y-axis.
#|    Bars are colored by the school year (ordered, factor)
#|    Plot illustrates the relationship between the three variables.
#| 

Gen_Growth_G5Math_Plot

```

## [Limitations? Do you do this or is that more for research?]{style="color:navy;"}

## [Appendix]{style="color:navy;"}

